### 2019-06-06 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Lots of small changes
- added a title/author etc in metadata
- cleaner small proof with an environment

\begin{proof}
    First, we consider the denominator in the expression of $\GLR(n)$ \eqref{eq:6:firstDefGLRT}, that is the $\sup$ on $\mu_0$.
    We have $\ell(X_1, \ldots, X_n ; \mu_0) = \prod_{i=1}^n \ell(X_i ; \mu_0)$ by independence of the observations $X_i$,
    and $\ell(X_i ; \mu_0) = \mu_0^{X_i} (1-\mu_0)^{1-X_i}$ for Bernoulli distributions.
    Therefore, taking the logarithm gives
    \begin{align*}
        \log\ell(X_1, \ldots, X_n ; \mu_0)
        &= \sum_{i=1}^n X_i \log(\mu_0) + (1-X_i) \log(1-\mu_0) \\
        &= \log(\mu_0) \times \left( \sum_{i=1}^n X_i \right) + \log(1-\mu_0) \times \left( n - \sum_{i=1}^n X_i \right) \\
        &= n \left( \widehat{\mu}_{1:n} \log(\mu_0) + (1 - \widehat{\mu}_{1:n}) \log(1-\mu_0) \right).
    \end{align*}
    %
    For a constant $a\in[0,1]$, let $h(x) := a \log(x) + (1-a) \log(1-x)$ on $(0,1)$,
    and we are trying to solve $\sup_{x\in[0,1]} h(x)$.
    If $a=0$ or $a=1$, $h$ is maximum at $x=a$.
    Now if $a\neq0$, As $h$ is of class $\cC^1$, we can just differentiate and find the root of its derivative:
    $h'(x) = a/x - (1-a)/(1-x)$, so $h'(x) = 0$ if and only if $x=a$.
    Thus in all cases, $\sup_{x\in[0,1]} h(x) = h(a)$.
    %
    Here, we have $a = \widehat{\mu}_{1:n}$, and thus we solve the $\sup_{\mu_0}$ optimization problem found in the denominator of the $\GLR(n)$ expression.
    By replacing $\mu_0 = \widehat{\mu}_{1:n}$, we obtained the following (unique) solution,
    \begin{equation}\label{eq:6:solutionPbOpt_for_1n}
        \sup_{\mu_0} \log\ell(X_1, \ldots, X_n ; \mu_0) = n \left( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \right).
    \end{equation}

@@

\begin{proof}
Lemma~\ref{lem:6:FalseAlarm} is presented for bounded distributions and is actually valid for any sub-Bernoulli distribution. It could also be presented for more general distributions satisfying
\begin{equation}
    \bE[e^{\lambda X} ] \leq e^{\phi_\mu(\lambda)} \ \ \text{ with } \ \ \mu=\bE[X]\label{def:6:subExpo},
\end{equation}
where $\phi_\mu(\lambda)$ is the $\mathrm{log}$ moment generating of some one-dimensional exponential family. The Bernoulli divergence $\kl(x,y)$ would be replaced by the corresponding divergence in that exponential family (which is the Kullback-Leibler divergence between two distributions of means $x$ and $y$).

Let us go back to the Bernoulli case with divergence $\kl(x,y)$.
We first observe that
\[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) = \inf_{\lambda \in [0,1]}\left[s \times \kl\left(\widehat{\mu}_{1:s},\lambda\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\lambda\right)\right].\]
Hence the probability of a false alarm occurring is upper bounded as
%
\begin{align*}
    \bP_{\mu_0}\left(T_{\delta} < \infty\right) & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) > \beta(n,\delta)\right)\\
    & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right) \\
% \end{align*}%FIXME uncomment this if needed
% \begin{align*}
    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists n > s:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right)\\
    & =   \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^*:  s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > \beta(s+r,\delta)\right),
\end{align*}
%
where $\widehat{\mu}_{s}$ and $\widehat{\mu}'_{r}$ are the empirical means of respectively $s$ and $r$ \iid{} observations with mean $\mu_0$ and distribution $\nu$, that are independent from the previous ones.
As $\nu$ is sub-Bernoulli, the conclusion follows from Lemma~\ref{lem:6:ConcFirst} below and from the definition of $\beta(n,\delta)$, if we denote $F(x)=\ln(1+\ln(x))$,
%
\begin{align*}
    &\bP_{\mu_0}\left(T_{\delta} < \infty\right) \\
    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 6F(s+r)+ 2 \cT\left(\frac{\ln(3(s+r)^{3/2}/\delta)}{2}\right)\right)\\
    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 3F(s)+3F(r) + 2 \cT\left(\frac{\ln(3s^{3/2}/\delta)}{2}\right)\right)
\end{align*}
%
And so we have $\bP_{\mu_0}\left(T_{\delta} < \infty\right) \leq \sum_{s =1}^{\infty} \frac{\delta}{3s^{3/2}} \leq \delta$.
%
\end{proof}

@@

    Let $\kl(\mu,\mu') = \KL(\nu^\mu,\nu^{\mu'})$ be the divergence function associated to that exponential family. Introducing the notation $\widehat{\mu}_s = \frac{1}{s}\sum\limits_{i=1}^s X_i$ and $\widehat{\mu}'_r = \frac{1}{r}\sum\limits_{i=1}^r Y_i$, it holds that for every $s,r \in \N^*$,
    \[\bP\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu\right) + r \, \kl\left(\widehat{\mu}_r',\mu'\right) > 3\ln(1+\ln(s)) + 3\ln(1+\ln(r)) + 2\cT\left(\frac{x}{2}\right)\right)\leq e^{-x},\]
    where $\cT$ is the function defined in \eqref{def:6:function_T}.
\end{leftbar}  % XXX leftbar lemmabar, comment if needed
\end{lemma}
\begin{proof}
    This Lemma~\ref{lem:6:ConcFirst} is proved in Appendix~\ref{app:6:proofConcFirst}.
\end{proof}

@@

\begin{lemma}\label{lem:6:Delay}
\begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
    Let be $\bP_{\mu_0,\mu_1,\tau}$ a probabilistic model under which $X_t \in [0,1]$, and $X_t$ has an average of $\mu_0$ for all $t \leq \tau$, and $\mu_1$ for all $t > \tau$, with $\mu_0 \neq \mu_1$,
    and let $\Delta = |\mu_0 - \mu_1|$.
    Then the B-GLRT test given in Definition~\ref{def:6:GLRDef} satisfies
    \begin{equation}
        \bP_{\mu_0,\mu_1,\tau} (\widehat{\tau}_{\delta} \geq \tau + u) \leq \exp\left( -\frac{2\tau u}{\tau + u}\left(\max\left[ 0, \Delta - \sqrt{\frac{\tau + u}{2\tau u} \beta(\tau + u,\delta)} \right]\right)^2 \right).
    \end{equation}
\end{leftbar}  % XXX leftbar lemmabar, comment if needed
\end{lemma}
%
\begin{proof}
    It is actually not used as such in the proofs of the regret upper bounds given below, because a similar but more specific result is used and proved in the proofs.
    For instance, see below in Section~\ref{par:6:controllingDelayInOneProof} in page~\pageref{par:6:controllingDelayInOneProof}.
    % \TODOL{We have to give the proof, or remove this result? Je ne vois pas le problème de le laisser là, ça formalise le résultat mais on a pas besoin de celui là, ça permet de bien formaliser ce qu'on entend par contrôler le délai, et dans la preuve plus bas on prouve ce qu'on souhaite.}
\end{proof}

@@

\begin{proof}
    We consider one arm $i\in[K]$, and when the \GLRklUCB{} algorithm is running,
    we consider two time steps $s\leq t\in\N^*$, chosen between two restart times for that arm $i$.
    Lines~$3$-$4$ state that $A_u = u \mod \lceil \frac{K}{\alpha} \rceil$
    if $u \mod \lceil \frac{K}{\alpha} \rceil \in [K]$,
    % (see Algorithm~\ref{algo:6:GLRklUCB} for details).
    thus we directly find
    \begin{align*}
        n_i(t) - n_i(s)
        & = \sum_{u=s+1}^{t} \indic(A_u = i) \\
        & \geq \sum_{u=s+1}^{t} \indic\left(A_u = i, A_u = u \mod \left\lceil \frac{K}{\alpha} \right\rceil\right) \\
        & \geq \sum_{u=s+1}^{t} \indic\left(i = u \mod \left\lceil \frac{K}{\alpha} \right\rceil\right) \\
        & = \bigl(t - (s+1) + 1\bigr) / \left\lceil \frac{K}{\alpha} \right\rceil
        \geq \left\lfloor \frac{\alpha}{K} (t - s) \right\rfloor.
    \end{align*}
    % Hence we have the result of Proposition~\ref{prop:6:EnoughSamples}.
\end{proof}

>>>>

\begin{smallproof}
    First, we consider the denominator in the expression of $\GLR(n)$ \eqref{eq:6:firstDefGLRT}, that is the $\sup$ on $\mu_0$.
    We have $\ell(X_1, \ldots, X_n ; \mu_0) = \prod_{i=1}^n \ell(X_i ; \mu_0)$ by independence of the observations $X_i$,
    and $\ell(X_i ; \mu_0) = \mu_0^{X_i} (1-\mu_0)^{1-X_i}$ for Bernoulli distributions.
    Therefore, taking the logarithm gives
    \begin{align*}
        \log\ell(X_1, \ldots, X_n ; \mu_0)
        &= \sum_{i=1}^n X_i \log(\mu_0) + (1-X_i) \log(1-\mu_0) \\
        &= \log(\mu_0) \times \left( \sum_{i=1}^n X_i \right) + \log(1-\mu_0) \times \left( n - \sum_{i=1}^n X_i \right) \\
        &= n \left( \widehat{\mu}_{1:n} \log(\mu_0) + (1 - \widehat{\mu}_{1:n}) \log(1-\mu_0) \right).
    \end{align*}
    %
    For a constant $a\in[0,1]$, let $h(x) := a \log(x) + (1-a) \log(1-x)$ on $(0,1)$,
    and we are trying to solve $\sup_{x\in[0,1]} h(x)$.
    If $a=0$ or $a=1$, $h$ is maximum at $x=a$.
    Now if $a\neq0$, As $h$ is of class $\cC^1$, we can just differentiate and find the root of its derivative:
    $h'(x) = a/x - (1-a)/(1-x)$, so $h'(x) = 0$ if and only if $x=a$.
    Thus in all cases, $\sup_{x\in[0,1]} h(x) = h(a)$.
    %
    Here, we have $a = \widehat{\mu}_{1:n}$, and thus we solve the $\sup_{\mu_0}$ optimization problem found in the denominator of the $\GLR(n)$ expression.
    By replacing $\mu_0 = \widehat{\mu}_{1:n}$, we obtained the following (unique) solution,
    \begin{equation}\label{eq:6:solutionPbOpt_for_1n}
        \sup_{\mu_0} \log\ell(X_1, \ldots, X_n ; \mu_0) = n \left( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \right).
    \end{equation}

@@

\begin{smallproof}
Lemma~\ref{lem:6:FalseAlarm} is presented for bounded distributions and is actually valid for any sub-Bernoulli distribution. It could also be presented for more general distributions satisfying
\begin{equation}
    \bE[e^{\lambda X} ] \leq e^{\phi_\mu(\lambda)} \ \ \text{ with } \ \ \mu=\bE[X]\label{def:6:subExpo},
\end{equation}
where $\phi_\mu(\lambda)$ is the $\mathrm{log}$ moment generating of some one-dimensional exponential family. The Bernoulli divergence $\kl(x,y)$ would be replaced by the corresponding divergence in that exponential family (which is the Kullback-Leibler divergence between two distributions of means $x$ and $y$).

Let us go back to the Bernoulli case with divergence $\kl(x,y)$.
We first observe that
\[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) = \inf_{\lambda \in [0,1]}\left[s \times \kl\left(\widehat{\mu}_{1:s},\lambda\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\lambda\right)\right].\]
Hence the probability of a false alarm occurring is upper bounded as
%
\begin{align*}
    \bP_{\mu_0}\left(T_{\delta} < \infty\right) & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) > \beta(n,\delta)\right)\\
    & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right) \\
% \end{align*}%FIXME uncomment this if needed
% \begin{align*}
    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists n > s:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right)\\
    & =   \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^*:  s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > \beta(s+r,\delta)\right),
\end{align*}
%
where $\widehat{\mu}_{s}$ and $\widehat{\mu}'_{r}$ are the empirical means of respectively $s$ and $r$ \iid{} observations with mean $\mu_0$ and distribution $\nu$, that are independent from the previous ones.
As $\nu$ is sub-Bernoulli, the conclusion follows from Lemma~\ref{lem:6:ConcFirst} below and from the definition of $\beta(n,\delta)$, if we denote $F(x)=\ln(1+\ln(x))$,
%
\begin{align*}
    &\bP_{\mu_0}\left(T_{\delta} < \infty\right) \\
    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 6F(s+r)+ 2 \cT\left(\frac{\ln(3(s+r)^{3/2}/\delta)}{2}\right)\right)\\
    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 3F(s)+3F(r) + 2 \cT\left(\frac{\ln(3s^{3/2}/\delta)}{2}\right)\right)
\end{align*}
%
And so we have $\bP_{\mu_0}\left(T_{\delta} < \infty\right) \leq \sum_{s =1}^{\infty} \frac{\delta}{3s^{3/2}} \leq \delta$.
%
\end{smallproof}

@@

    Let $\kl(\mu,\mu') = \KL(\nu^\mu,\nu^{\mu'})$ be the divergence function associated to that exponential family. Introducing the notation $\widehat{\mu}_s = \frac{1}{s}\sum\limits_{i=1}^s X_i$ and $\widehat{\mu}'_r = \frac{1}{r}\sum\limits_{i=1}^r Y_i$, it holds that for every $s,r \in \N^*$,
    \[\bP\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu\right) + r \, \kl\left(\widehat{\mu}_r',\mu'\right) > 3\ln(1+\ln(s)) + 3\ln(1+\ln(r)) + 2\cT\left(\frac{x}{2}\right)\right)\leq e^{-x},\]
    where $\cT$ is the function defined in \eqref{def:6:function_T}.
\end{leftbar}  % XXX leftbar lemmabar, comment if needed
\end{lemma}
\begin{smallproof}
    This Lemma~\ref{lem:6:ConcFirst} is proved in Appendix~\ref{app:6:proofConcFirst}.
\end{smallproof}

@@

\begin{lemma}\label{lem:6:Delay}
\begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
    Let be $\bP_{\mu_0,\mu_1,\tau}$ a probabilistic model under which $X_t \in [0,1]$, and $X_t$ has an average of $\mu_0$ for all $t \leq \tau$, and $\mu_1$ for all $t > \tau$, with $\mu_0 \neq \mu_1$,
    and let $\Delta = |\mu_0 - \mu_1|$.
    Then the B-GLRT test given in Definition~\ref{def:6:GLRDef} satisfies
    \begin{equation}
        \bP_{\mu_0,\mu_1,\tau} (\widehat{\tau}_{\delta} \geq \tau + u) \leq \exp\left( -\frac{2\tau u}{\tau + u}\left(\max\left[ 0, \Delta - \sqrt{\frac{\tau + u}{2\tau u} \beta(\tau + u,\delta)} \right]\right)^2 \right).
    \end{equation}
\end{leftbar}  % XXX leftbar lemmabar, comment if needed
\end{lemma}
%
\begin{smallproof}
    It is actually not used as such in the proofs of the regret upper bounds given below, because a similar but more specific result is used and proved in the proofs.
    For instance, see below in Section~\ref{par:6:controllingDelayInOneProof} in page~\pageref{par:6:controllingDelayInOneProof}.
    % \TODOL{We have to give the proof, or remove this result? Je ne vois pas le problème de le laisser là, ça formalise le résultat mais on a pas besoin de celui là, ça permet de bien formaliser ce qu'on entend par contrôler le délai, et dans la preuve plus bas on prouve ce qu'on souhaite.}
\end{smallproof}

@@

\begin{smallproof}
    We consider one arm $i\in[K]$, and when the \GLRklUCB{} algorithm is running,
    we consider two time steps $s\leq t\in\N^*$, chosen between two restart times for that arm $i$.
    Lines~$3$-$4$ state that $A_u = u \mod \lceil \frac{K}{\alpha} \rceil$
    if $u \mod \lceil \frac{K}{\alpha} \rceil \in [K]$,
    % (see Algorithm~\ref{algo:6:GLRklUCB} for details).
    thus we directly find
    \begin{align*}
        n_i(t) - n_i(s)
        & = \sum_{u=s+1}^{t} \indic(A_u = i) \\
        & \geq \sum_{u=s+1}^{t} \indic\left(A_u = i, A_u = u \mod \left\lceil \frac{K}{\alpha} \right\rceil\right) \\
        & \geq \sum_{u=s+1}^{t} \indic\left(i = u \mod \left\lceil \frac{K}{\alpha} \right\rceil\right) \\
        & = \bigl(t - (s+1) + 1\bigr) / \left\lceil \frac{K}{\alpha} \right\rceil
        \geq \left\lfloor \frac{\alpha}{K} (t - s) \right\rfloor.
    \end{align*}
    % Hence we have the result of Proposition~\ref{prop:6:EnoughSamples}.
\end{smallproof}
